---
title: "assignment"
author: "Gavin Graf"
date: "2024-04-23"
output: html_document
---

I had issues using the forceseqlevelstyle of Ensembl, so I manually made the seqlevels compatible. This made it possible to use the signal2matrix function, but I was unable to solve the errors from the functions getnormfactors and plotenrichedheatmaps. From what I understand the objects in the tracks need to be granges objects, which they are. The plotenrichedheatmaps error is due to the size of the termporary image for rasterization which I tried to limit, but unsuccessfully.

```{r packages}
suppressPackageStartupMessages({
  library(epiwraps)
  library(AnnotationHub)
  library(MotifDb)
  library(memes)
  library(universalmotif)
  library(ensembldb)
  library(ggplot2)
  library(rtracklayer)
})
```

```{r bigwig files}

download.file("https://ethz-ins.org/content/w7/atac.chr19.bam", "atac.chr19.bam", mode = "wb")
download.file("https://ethz-ins.org/content/w7/atac.chr19.bam.bai", "atac.chr19.bam.bai", mode = "wb")

bam <- "atac.chr19.bam"


bam2bw(bam, output_bw = "NF_cuts.bw", paired=TRUE, binWidth=1L, type="ends", extend=2L, minFragLength=30, 
       maxFragLength=120, shift=c(4L,-5L), forceSeqlevelsStyle = "UCSC")

bam2bw(bam, output_bw = "N_centers.bw", paired=TRUE, binWidth=5L, minFragLength=140, shift=c(4L,-5L), 
       maxFragLength=220, type="center", extend=10L, forceSeqlevelsStyle = "UCSC")

bwNF <- "NF_cuts.bw"

bwNC <- "N_centers.bw"

bwNFgr <- import(bwNF)
bwNCgr <- import(bwNC)
chr <- "chr19" 
bwNFgr19 <- subset(bwNFgr, seqnames(bwNFgr) == chr)
bwNCgr19 <- subset(bwNCgr, seqnames(bwNCgr) == chr)
seqlevels(bwNFgr19) <- chr
seqlevels(bwNCgr19) <- chr
seqlevels(bwNFgr19) <- sub("^chr", "", seqlevels(bwNFgr19))
seqlevels(bwNCgr19) <- sub("^chr", "", seqlevels(bwNCgr19))
```

```{r motif}
KLF4 <- MotifDb::query(MotifDb, c("KLF4","Mus"))[[1]]
MAZ <- MotifDb::query(MotifDb, c("MAZ","Mus"))[[1]]

KLF4c <- convert_motifs(KLF4, class="TFBSTools-PFMatrix")
MAZc <- convert_motifs(MAZ, class="TFBSTools-PFMatrix")

ah <- AnnotationHub()
genome <- ah[["AH68356"]]
chr19 <- import(genome)["19"]



KLF4moi <- motifmatchr::matchMotifs(KLF4c, chr19, out="positions", p.cutoff=1e-5)[[1]]
MAZmoi <- motifmatchr::matchMotifs(MAZc, chr19, out="positions", p.cutoff=1e-5)[[1]]

KLF4moi2 <- as(setNames(KLF4moi,names(chr19)), "GRanges")
MAZmoi2 <- as(setNames(MAZmoi,names(chr19)), "GRanges")


```

```{r plot}

tracks <- c("NF Cuts"= bwNFgr19, "Nucleosome Centers"= bwNCgr19)

smKLF4 <- signal2Matrix(tracks, KLF4moi2, w=3, extend=300)
smMAZ <- signal2Matrix(tracks, MAZmoi2, w=3, extend=300)


nf <- getNormFactors(tracks, useSeqLevels = "19", nwind=5000L)

smKLF4 <- renormalizeSignalMatrices(sm, scaleFactors = nf)
smMAZ <- renormalizeSignalMatrices(MAZ, scaleFactors = nf)


ht_opt(raster_temp_image_max_width = 200, raster_temp_image_max_height = 2000)¨

plotEnrichedHeatmaps(smKLF4, trim=0.95)
plotEnrichedHeatmaps(smMAZ, trim=0.95)




```